---
title: "Declaring and Diagnosing Research Designs"
author: "Graeme Blair, Jasper Cooper, Alexander Coppock, Macartan Humphreys"
output:
  bookdown::gitbook:
    includes:
      in_header: [hypothesis.html]
    config:
      download: []
      toc:
        after: |
          <li><a href="https://declaredesign.org" target="blank">DeclareDesign</a></li>
        before: |
          <li><a href="./">Declaring and Diagnosing Research Designs</a></li>
        collapse: section
      sharing:
        all: false
        github: false
        facebook: false
        twitter: false
        google: false
        weibo: false
        instapper: false
    css: style.css
    split_by: section
bibliography: [bib.bib, packages.bib]
description: "Declaring and diagnosing research designs in the model, inquiry, data strategy, analysis strategy (MIDA) framework"
documentclass: book
fontfamily: mathpazo
fontsize: 11pt
github-repo: rstudio/rdb
link-citations: yes
site: bookdown::bookdown_site
biblio-style: apalike
---


```{r include=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


```{r, echo=FALSE, warning=FALSE}
library(DeclareDesign)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(forcats)
library(knitr)

set.seed(42)

rerun_templates <- FALSE
sims <- 10000

intro_path   <- "01_Part_A_MIDA"
designs_path <- "02_Part_B_Designs"
outro_path   <- "03_Part_C_Principles"
principles_path <- "03_Part_C_Principles"

library(ggplot2)

pro_con_colors <- c("#C67800", "#205C8A")

#colorfunction <- colorRampPalette(pro_con_colors)
#pred_cols <- colorfunction(4)

dd_theme <-
  function() {
    theme_bw() +
      theme(
        axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_line(color = '#eeeeee'),
        strip.background = element_blank(),
        legend.position = "bottom",
        text = element_text(family = "Palatino"))
  }

#p <- ggplot(mtcars, aes(wt, mpg))
#p + geom_point() + dd_theme()

```

# What we are at {-}

We have big hopes for this book. We hope to promote a new, comprehensive way of thinking about research designs in the social sciences. We hope this way of thinking will make research designs more transparent and more robust. But we also hope it will make research design easier, easier to produce good designs, but also easier to share designs and build off of the designs that others have developed.

The core idea is to start think of a design as an object that can be interrogated. The design encodes your beliefs about the world, it describes your questions, and it lays out  how you go about answering those questions, in terms both of what data you use and how you use it. A key idea is that all of these features can be provided in code and if done right the information provided is enough to be able to simulate a run of the design. being able to simulate a design puts a reearcher in a powerful position as you can then start assessing the conditions under which a design perfoms well or badly.



