```{r,echo=FALSE,eval=FALSE}
tempR <- tempfile(fileext = ".R")
library(knitr)
purl("../index.Rmd", output=tempR)
source(tempR)
unlink(tempR)
```

Does the way in which one causal factor affects an outcome change when some other causal factor is present? This very common question is often studied in experiments using the two-by-two factorial design. Christopher Blattman, Julian C. Jamison and Margaret Sheridan, for example, sought to test the theory that giving cash to criminally-engaged men in Uganda would produce a bigger reduction in their subsequent criminal behavior if it was coupled with cognitive behavioral therapy. The authors reasoned that cash windfalls would be especially likely to shift those with high business ability to non-criminal careers where those abilities were comparatively more valuable, and that therapy would increase business ability. The authors find that while cash and therapy had moderate affects on their own, their combination caused a substantial and enduring reduction in crime and violence. The authors infer that the effect of cash ($Z_1$) on crime ($Y$) is a function of behavioral therapy ($Z_2$) because the interaction of $Z_1$ and $Z_2$ on $Y$ is statistically significantly different from zero. 

Before conducting studies of this kind, researchers often want to increase their confidence in their ability to detect significant interactions when the effect of $Z_1$ on $Y$ really is a function of $Z_2$. How many subjects are needed to raise the probability of correctly inferring that there is an interaction 80\%?

This is a surprisingly difficult question to answer, because answering it accurately depends upon the specific model of the potential outcomes the researcher has in mind. As a result, many researchers use rules of thumb that may lead to systematic over- or under-confidence. Using the MIDA framework and design simulation, however, we can provide a flexible answer to this question that does not rely on rules of thumb.

### MIDA

- **M**odel: We specify $Z_1$ and $Z_2$ do not have any effect on the outcome when only one of the causal agents becomes present. When both are present the combination of the causal factors produces an increase of 1/10th of a standard deviation in the outcome.

- **I**nquiry: We can express the effect of $Z_1$ when $Z_2$ is present as $\tau_{Z_1 \mid Z_2} = E[(Y \mid Z_1 = 1, Z_2 = 1) - (Y \mid Z_1 = 0, Z_2 = 1)]$, and the effect of $Z_1$ when $Z_2$ is absent as $\tau_{Z_1 \mid \neg Z_2} = E[(Y \mid Z_1 = 1, Z_2 = 0) - (Y \mid Z_1 = 0, Z_2 = 0)]$. Thus, our estimand is $\tau_{Z_1 \mid Z_2} - \tau_{Z_1 \mid \neg Z_2}$: the difference in the effect of $Z_1$ induced by moving $Z_2$ from 0 to 1.

- **D**ata strategy: We randomly assign an equal number of subjects to one of four conditions. In the first both causal factors are absent, in the second and third only $Z_1$ or $Z_2$ is present, respectively, and in the fourth both are present. 

- **A**nswer strategy: We estimate the interaction effect using a linear regression model that focuses on the coefficient on the $Z_1 \times Z_2$ term.

### Declaration

Our `factorial_template` function takes as arguments the total sample size (`N`), the effect of $Z_1$ (`beta_1`), the effect of $Z_2$ (`beta_2`), and the effect of $Z_1 \times Z_2$ (`beta_3`). 

```{r,echo = TRUE, eval = rerun_templates}
factorial_template <-
  function(N, beta_1 = 0, beta_2 = 0, beta_3) {
    population <- declare_population(noise = rnorm(N), N = N)
    potential_outcomes <-
      declare_potential_outcomes(
        Y_Z_0 = noise,
        Y_Z_1 = beta_1 + noise,
        Y_Z_2 = beta_2 + noise,
        Y_Z_3 = beta_1 + beta_2 + beta_3 + noise
      )
    assignment_factorial <-
      declare_assignment(condition_names = 0:3)

    estimand <-
      declare_estimand(interaction_effect = mean((Y_Z_3 - Y_Z_2) - (Y_Z_1 - Y_Z_0)))

    estimator <- declare_estimator(
      formula = Y ~ Z1 + Z2 + Z1 * Z2,
      model = lm_robust,
      coefficient_name = "Z1:Z2",
      estimand = estimand
    )
    declare_design(
      population,
      potential_outcomes,
      assignment_factorial,
      mutate(Z1 = as.numeric(Z %in% c(1, 3)),
             Z2 = as.numeric(Z %in% c(2, 3))),
      reveal_outcomes,
      estimator,
      estimand
    )
  }
```

```{r, eval=FALSE}
designs <- quick_design(factorial_template,
                        N = seq(200, 2000, 200),
                        beta_3 = .25)
diagnoses <- diagnose_design(designs, sims = 1000, bootstrap = FALSE)
```

```{r, echo=FALSE, eval = !rerun_templates}
diagnoses <- readRDS("../shiny_applications/factorial/data/factorial_forshiny.RDS")
```

### Takeaways

```{r, echo=FALSE}
diagnosand_plot <- diagnoses$diagnosands %>%
  filter(beta_3 == .25) %>% 
  ggplot(aes(N, power)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.8, color = "red") +
  xlab("Number of Subjects") +
  scale_y_continuous("Power to Detect a 0.25 SD Interaction",breaks = c(seq(0,1,.25),8), limits = c(0,1)) +
  ggtitle("2x2 Factorial Experiment", "Power Analysis") +
  theme_bw()

if (knitr::opts_knit$get("rmarkdown.pandoc.to") != "html") {
  diagnosand_plot
}
```

<iframe src="https://acoppock.shinyapps.io/factorial/" style="border: none; width: 100%; height: 600px"></iframe>

### Exercises

1. Alter the template so that outcomes are binary instead of normally distibuted. What is the expected standard error for the interaction term for a sample size of 1000? Discuss the implications of your diagnosis for practice.

2. Heterogeneous shifts by making effects noise
